Adopting agile in the large: motivations, success factors and challenges

### 2014-05-16 journal/article.tex
Editing challenges

@@

\begin{figure}[!b]
  \begin{center}
    \includegraphics{graphics/challenges_summary.pdf}
    \caption{Challenges in agile transformation}
    \label{fig:challenges_summary}
  \end{center}
\end{figure}

@@

It is clear that changing the way of working will create some costs for the
organization. If there is no intention to allow resources for the transformation
problems will ensue. Typical examples on this was lack of investment in
training, and that people were kept allocated to tasks from the original
arrangement. In some cases the transformation was impeded when work spaces could
not be updated.

@@

Lack of resources allocated for supporting a transformation was an evident
problem. Resources were most often short for training and coaching, but
underbudgeting these activities may be a serious mistake.
Hajjdiab [P20] describes how reluctance of management to invest in training
caused teams to be ill prepared for the agile transformation effort, and
eventually resulted in ending the use of agile methods. As a less dramatic
outcome of shortened training was lower motivation [P45].

Proper coaching was a particular problem for very large organizations. In cases
where the number of teams in need of coaching was large the demand often
exceeded the capacity of available coaches [P1, P6, P23, P49]. Lack of coaching
was also attributed as a reason why the success of pilot teams could not be
repeated when agile was adopted more widely [P9, P45].
Silva and Doss [P47] described that when it was hard to fill the coaching
positions people less seasoned in agile were appointed as coaches, which risked
that agile practices would not be taught correctly.

@@

A condition hindering the transformation was that old commitments still needed
to be tended to even though the organization was being restructured.
Some organizations started their agile journey in a state where people were
overcommitted, but it was later realized that overloaded people will not be able
to learn new ways of working and changing their behavior [P37, P42]. Cowan [P11]
describes how time was allocated for tending to old commitments. However, even
when there was preparation for extra work people with specific expertise became
overloaded. After time it was realized that overcommitment must be reduced by
allowing to shift excess work for later time [P11]. In another case management
forced people to remain committed to firm deadlines, even midst in the
organizational transformation. The engagement in old commitments and tasks
resulted in ignoring new practices introduced by agile, and eventually the agile
method was breaking down [P21]. Spayd [P49] also describes how senior management 
pressed teams to deliver what the customer had been promised, regardless of what
the current velocity of development predicted.

@@

A challenge most organizations came across was that an agile model was difficult
implement. An experienced software team may do well in training, but when the
time comes to apply agile techniques in practice the team may get lost [P21].
Agile was also thought being a solution to all problems [P9], success was
declared prematurely [P49], and expectations created by successful early
experiments were not fullfilled [P38]. Disapointment was granted when these
hopes were not fullfilled, but instead implementing the agile model was found
to be hard and time consuming. The difficulty in implementing agile is evident
in misconceptions of the agile model, when by-the-book implementations prove to
be insufficient, and when the agile methods were applied badly to the
organization.

@@

On a team level there were several examples of misunderstanding the agile model.
Cases describe how agile was seen as freedom to hack without documentation
[P26], that development could be done without design tasks [P37], and that agile
meant that everyone should become a generalist [P42]. Schatz and Abdelshafi
describe how teams presented unfinished work at reviews and ignored the
principle of showing only completed increments, which resulted in a backlog of
bugs [P44]. In one case team members had perceived the introduction of agile as
a means to squeeze out more efficiency [P45]. In addition, the flatter
organization was seen as fewer career opportunities, and pushed team members to
internal competition for visibility [P45].
Still one misunderstanding of agile was due to perceiving it only through the
tools used, such as project management software [P48]. Superficial focusing only
on tools themselves and not the reasons behind their use led to frustration
[P42].

@@

@@

Even if there was no misunderstandings the correct agile model was found hard to
implement. Several cases describe how an agile model was hard to learn from
literature [P21, P27], especially when it comes to implementing it on a large
scale [P13]. As Beavers [P5] writes: ``There simply was not a manual or document
where we could find easy answers on how to do things''.
Schnitter and Mackert [P45] report that theoretical considerations on how to
scale up the agile methos were not good enough, and that product managers and
architects struggled when several Scrum teams were working concurrently.
Another case concluded that all practices prescribd in XP did not fit enterprise
scale development, and some practices required customization [P49]. The model of
agile should not be prescribed as a check list, as it would impose a single
model in an organization with several individual teams and projects [P41].

% Check this paragraph
Further indications were given that the advice given in literature will not
ferry all the way. Benefield [P6] describes how it was dificult finding a
balance between prescribing a by-the-book implementation which may put people
off, and giving too much freedom in the agile methods which may weaken core
practices.
The reality where the practices needed to be applied was described as
messy in comparison to the ideal circumstances presented in literature, which
made it difficult to choose an initial model and gain buy-in for it [P10].
Difficulties in choosing the model to start with were also due to variances
perceived in the agile approaches suggested in literature [P30].

@@

Many cases reported that introducing agile practices soon yielded improvements
on the team level, but as a part of a large transformation teams needed to
change their external behavior also.
The initial success was followed by challenges when teams needed to work with
other teams and as parts of the larger surrounding orgaization [P13, P17, P24].
Introducing agile had created flexibility on a team level, but it could not be
fully utilized as the surrounding organization was not responsive enough [P26].
There were dependencies between teams in the agile model as well, and the
dependencies made managing development difficult [P9, P10].

@@

Some organizations attempted to initially create a model where teams operate
autonomously, which is well in line with agile principles. However, a number of
issues arose from independence. For instance, teams needed to find a balance
between the their own goals and broader goals of the entire organization, but
some teams chose to focus only on their own goal [P7]. Independent teams that
did not respect the larger context they were working in obstructed coordination
[P13]. Coordination was also difficult when teams were working independently for
different customers but the applications being built were interdependent [P29].
One case reported that allowing teams to have different sprint durations created
delays in delivery [P51].

@@

Traditionally development is preceeded by requirements engineering and succeeded
by quality assurance (QA) activities. Agile methods suggest that essential parts
of these activities would be performed within the development team. The effort
required for managing requirements and QA for large systems may be too large for
individual teams to handle.

@@

In agile methods requirements engineering is usually closely tied to the
development team. However, large development projects will demand additional
high level requirements management.
This is apparent in cases where complexity of the product will require
additional layers of product management [P45], requirements are created by
several stakehoder groups and development teams can not be in contact with all
of them [P39], or teams do not have access to stakeholders due to a distributed
seetting [P24].
As an integral part of the software development lifecycle, requirements
management will be affected by the new agile process. In large organizations
product and requirements management was initially done in a traditional way, but
it was contrasted by the agile process that required iteratively supplying small
requirements to development [P51]. The agile method had to be extended to
include a process on how to get large requirements into team backlogs [P31].

Requirements were in some cases first defined on a high level in marketing
requirements documents (MRD) [P1, P31] or functional specifications [P33]. These
high level requirements needed to be elaborated to be used by agile development
teams. The requirements breakdown in it self was found to be difficult [P1,
P33], and also when to do the breakdown and to which detail level [P17].
Product managers and business analysts were struggling in creating high level
requirements [P33, P37], and teams were struggling in breaking them down to an
estimable size [P33]. Smits and Rilliet [P48] describe how high level product
management delivered requirements in too large chunks resulting in development
teams spending large amounts of time on defining suitaby sized stories. In one
case the teams were overwhelmed by reprioritization and requirements bartering
when stakeholders learned that requirements could be rearranged flexibly [P31].

@@

@@

High level requirements work and estimation was problematic as there was a gap
between short and long term planning [P9, P10]. Typical agile backlogs would
give only short term visibility, so organization specific practices for business
to do long term planning had to be created [P26, P31]. Several cases described
that care had to be taken to avoid long term planning becoming a prescriptive
practice. In order to preserve the agility of development a schedule driven
approach was avoided [P10], sensitivity was used when considering setting
milestones [P13], and striving for more accurate estimates was not let to become
an excuse for gathering requirements up front [P38].

@@

In an enterprise setting quality assurance tasks will require effort beyond
the general agile convention of doing testing within a crossfunctional team. In
large development efforts the agile model may need to be extended to accomodate
additional testing [P31]. Organizations may have existing separate QA teams that
must be coordinated to work with development teams [P17, P30, P48]. Integrating
enterprise scale QA will pose some additional challenges to agile methods.

Full quality assurance of a system requires special testing such as performance,
load, and memory testing, but agile methods lacks focus on them [P17]. These
testing activities can not be done within boundaries of user stories, and
require more resources than teams can spare [P31]. As testing tasks overlap team
boundaries it is sensible having a separate teams for the tasks, but
coordination practice between specialized QA teams and development teams needs
to be defined [P28, P48].

@@

Still one challenge in QA was the relationship to requirements engineering. The
QA functions were struggling if there was problems such as ambiguity or
ineffective breakdown of large requirements [P48]. In a waterfall model an
extended time was reserved for QA at the end of projects, which allowed to
retroactively resolve ambiguities in requirements [P5]. However, as agile
required QA work to be done incrementally with development problems with
ambiguity in requirements became an impediment to QA [P5].

@@

>>>>

* People centricity -- agile is not for squeezing efficiency -- agile is not for
  fewer career opportunities in a flatter organization.

* Agile is a mindset -- it can not be studied from a book -- need of coaching

@@

\begin{figure}[!b]
  \begin{center}
    \includegraphics{graphics/challenges_summary.pdf}
    \caption{Overview of challenges in agile transformation}
    \label{fig:challenges_summary}
  \end{center}
\end{figure}

@@

It is clear that changing the way of working will create some costs for the
organization. Problems will ensue if there is no intention to allow resources
for the transformation. Examples on this was lack of investment in training, and
keeping people allocated to tasks from the old arrangement. In some cases the
transformation was impeded when work spaces could not be updated.

@@

Lack of activities supporting a transformation was an evident problem. Training
and coaching were often short of resources, but underbudgeting these activities
may seriously hamper the transformation.
Hajjdiab [P20] describes how reluctance of management to invest in training
caused teams to be ill prepared for the agile transformation effort, and
eventually resulted in ending the use of agile methods. As a less dramatic
outcome of shortened training was lower motivation [P45].

Arranging proper coaching was a particular problem in many organizations. It is
critical to coach teams in their real work environment as a mindset change is
difficult to achieve by only attending training sessions. In large organizations
where numerous teams needed to be coached the demand often exceeded the capacity
of available coaches [P1, P6, P23, P49]. Lack of coaching was also attributed as
a reason why the success of pilot teams could not be repeated when agile was
adopted more widely [P9, P45].
Silva and Doss [P47] described that when it was hard to fill the coaching
positions people less seasoned in agile were appointed as coaches, which risked
that agile practices would not be taught correctly.

@@

Even though the organization was in a state of transformation the workload of
teams was not always adjusted to facilitate the process.
Some organizations started their agile journey in a state where people were
overcommitted, but it was later realized that when people are overloaded they
will not be able to changing their behavior and learn new ways of working [P37,
P42]. Cowan [P11] describes how time was allocated for tending to old
commitments. However, even when there was preparation for extra work people with
specific expertise became overloaded. After time it was realized that
overcommitment must be reduced by allowing to shift excess work for later time
[P11]. In another case management forced people to remain committed to firm
deadlines, even midst in the organizational transformation [P21]. The engagement
in old commitments and tasks resulted in ignoring new practices introduced by
agile, and eventually the agile method was breaking down [P21]. Spayd [P49] also
describes how senior management pressed teams to deliver what the customer had
been promised, regardless of what the current velocity of development predicted.
For one part the pressure to deliver interfered with the transformation, but it
was also forcing a change in the old way of working [P49].

@@

A challenge most organizations came across was that the agile model was
difficult to implement. An experienced software team may do well in training,
but when the time comes to apply agile techniques in practice the team may get
lost [P21]. The difficulty in implementing agile is evident in misconceptions of
the agile model, when by-the-book implementations prove to be insufficient, and
when the agile methods were applied badly to the organization.

@@

On a team level there were several examples of misunderstanding agile
development. Cases describe how agile was seen as freedom to hack without
documentation [P26], that development could be done without design tasks [P37],
and that agile meant that everyone should become a generalist [P42]. Schatz and
Abdelshafi describe how teams presented unfinished work at reviews and ignored
the principle of showing only completed increments, which resulted in a backlog
of bugs [P44]. In one case team members had perceived the introduction of agile
as a means to squeeze out more efficiency [P45]. In addition, the flatter
organization was seen as fewer career opportunities, and pushed team members to
internal competition for visibility [P45].
Still one misunderstanding of agile was due to viewing it only through the tools
used, such as project management software [P48]. Superficial focusing only on
tools themselves and not the reasons behind their use led to frustration [P42].

@@

As a final example of misconceptions, some organizations considered agile being
a solution to all problems [P9]. Success was declared prematurely [P49] but
expectations created by successful early experiments were not fullfilled [P38].
Disapointment was granted when these hopes were not fullfilled, and instead
implementing the agile model was found to be hard and time consuming.

@@

Even if there was no misunderstandings the correct agile model was hard to
implement. Several cases describe how an agile model was hard to learn from
literature [P21, P27], especially when it comes to implementing it on a large
scale [P13]. As Beavers [P5] writes: ``There simply was not a manual or document
where we could find easy answers on how to do things''.
Schnitter and Mackert [P45] report that theoretical considerations on how to
scale up the agile methos were not good enough, and that product managers and
architects struggled when several Scrum teams were working concurrently.
Another case concluded that all practices suggested by XP did not fit enterprise
scale development, and some practices required customization [P49]. The model of
agile should not be prescribed in a strict way, as it would impose a single
model in an organization with several individual teams and projects [P41].

% Check this paragraph
Further indications were given that the advice from literature will not ferry
all the way. Benefield [P6] describes how it was dificult finding a balance
between prescribing a by-the-book implementation which may put people off, and
giving too much freedom in the agile methods which may weaken core practices.
The reality where the practices needed to be applied was described as messy in
comparison to the ideal circumstances presented in literature [P10]. Because of
this it was difficult to choose a model to start with and gain buy-in for [P10].
Difficulties in choosing a initial model were also due to variances perceived in
the agile approaches suggested in literature [P30].

@@

Many cases reported that introducing agile practices soon yielded improvements
on the team level, but as a part of a large transformation teams also needed to
change their external behavior.
The initial success was followed by challenges when teams needed to work with
other teams and as parts of the larger surrounding orgaization [P13, P17, P24].
Introducing agile had created flexibility on a team level, but it could not be
fully utilized as the surrounding organization was not responsive enough [P26].
There were dependencies between teams in the agile model as well, and the
dependencies made managing development difficult [P9, P10].

@@

Some organizations attempted to initially create a model where teams operate
autonomously, which is well in line with agile principles. However, a number of
issues arose from independence. For instance, teams needed to find a balance
between the their own goals and broader goals of the entire organization, but
some teams chose to focus only on their own goal [P7]. Independent teams that
did not respect the larger context obstructed coordination [P13]. Coordination
was also difficult when teams were working independently for different customers
but the applications being built were interdependent [P29]. One case reported
that even allowing teams to have different sprint durations created delays in
delivery [P51].

@@

Traditionally development is preceeded by requirements engineering and succeeded
by quality assurance (QA) activities. Agile methods suggest that essential parts
of these activities would be performed within the development team. However,
when developming large systems the effort required for managing requirements and
QA may be too heavy for individual teams to handle.

@@

In agile methods requirements engineering is usually closely tied to the
development team. However, large development projects will demand additional
high level requirements management. This is apparent in cases where complexity
of the product will require additional layers of product management [P45],
requirements are created by several stakehoder groups and development teams can
not be in contact with all of them [P39], or teams do not have access to
stakeholders due to a distributed seetting [P24]. Agile methods need to be
extended to include a process on how to get large requirements into team
backlogs [P31].

% As an integral part of the software development lifecycle, requirements
% management will be affected by the new agile process. In large organizations
% product and requirements management was initially done in a traditional way, but
% it was contrasted by the agile process that required iteratively supplying small
% requirements to development [P51].

@@

Requirements were in some cases first defined on a high level in marketing
requirements documents (MRD) [P1, P31] or functional specifications [P33].
These high level requirements needed to be elaborated to be used by agile
development teams. The requirements breakdown in it self was found to be
difficult [P1, P33], and also when to do the breakdown and to which detail level
[P17].
Product managers and business analysts were struggling in creating high level
requirements [P33, P37], and teams were struggling in breaking them down to an
estimable size [P33]. Smits and Rilliet [P48] describe how high level product
management delivered requirements in too large chunks resulting in development
teams spending large amounts of time on defining suitaby sized stories.
% ?
In one case the teams were overwhelmed by reprioritization and requirements
bartering when stakeholders learned that requirements could be rearranged
flexibly [P31].

@@

High level requirements work and estimation was problematic as there was a gap
between short and long term planning [P9, P10]. Typical agile backlogs would
give only short term visibility, so organization specific practices for long
term planning had to be created [P26, P31]. Several cases described that care
had to be taken to avoid long term planning becoming a prescriptive practice. In
order to preserve the agility of development a schedule driven approach was
avoided [P10], sensitivity was used when considering setting milestones [P13],
and striving for more accurate estimates was not let to become an excuse for
gathering requirements up front [P38].

@@

In an enterprise setting quality assurance tasks will require effort beyond the
general agile convention of doing testing within a crossfunctional team.
Similarly as with requirements engineering, the agile model may need to be
extended to accomodate additional testing activities [P31]. Organizations may
have existing separate QA teams that must be coordinated to work with
development teams [P17, P30, P48]. Integrating enterprise scale QA will pose
some additional challenges to agile methods.

Full quality assurance of a system requires special testing such as performance,
load, and memory testing, but agile methods lacks focus on them [P17]. These
testing activities can not be done within boundaries of user stories, and
require more resources than teams can spare [P31]. When testing tasks overlap
team boundaries it is sensible having a separate teams for the tasks, but
coordination practice between specialized QA teams and development teams needs
to be defined [P28, P48].

@@

Still one challenge in QA was the relationship to requirements engineering. The
QA functions were struggling if there was problems such as ambiguity or
ineffective breakdown of large requirements [P48]. In a waterfall model an
extended time was reserved for QA at the end of projects, which allowed to
retroactively resolve ambiguities in requirements [P5]. However, ambiguity
became an impediment to QA when development worked in short increments and
there was no extended period for stabilization [P5].

@@

* Have flexibility for different teams in the agile model (agile found
  hard [P41]) vs. conformity to a single model.
